Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-07-02 paper/ms.tex
update ms

\newcommand{\ntwoyear}{27} % TODO: update
\newcommand{\ntwoyearturnedoff}{5} % TODO: update

@@

%\title{Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS}

\title{Corotating Dust Clumps Around Adolescent Low-Mass Stars: Four Years of TESS}

%\title{Complex Quasiperiodic Variables From Four Years of TESS}

@@

% primary authors
\author[0000-0002-0514-5538]{Luke~G.~Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{\caltech}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust or gas clumps in orbit at the
  Keplerian corotation radius.  Here, we report new CQVs discovered in
  TESS data collected between July~2018 and Sep~2022.
  Our search of \nstarssearched\ K- and M-dwarfs
  with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\ gold-standard CQVs.
  Most of these discoveries are new, and they include the brightest
  and closest examples of this class of object known ($T$$\approx$9.5;
  $d$$\approx$20\,pc), as well as the most massive
  ($\approx$0.65\,M$_\odot$).  A few objects are outliers among
  outliers; LP 12-502 for instance shows a ``dip complex'' with a
  period and total duration that are fixed over more than 1{,}000 cycles,
  but which in detail shows anywhere from four to eight local minima
  per cycle.  LP-502 at times demonstrated drastic shape changes over
  less than one cycle, and also displayed distinct superposed periods
  simultaneously.  Broadly speaking, we find that none of the CQVs
  maintain a fixed light curve shape over timescales of more than a
  few hundred cycles, and we revisit the arguments for why transient
  corotating material is the most likely explanation.  In the
  future, we expect that our sample will facilitate modeling and
  observational efforts aimed at understanding these objects, and
  connecting them to the broader contexts of star, disk, and exoplanet
  evolution.
%  For instance, based on the rate of dip evolution, we
%  estimate that the incident dust flux corresponds to of order an
%  asteroid belt's worth of mass accreting onto these stars over $10^8$
%  years.
\end{abstract}

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.  More exotic forms of optical variability
relevant to this work include transiting exocomets (e.g. $\beta$~Pic;
\citealt{2019A&A...625L..13Z}) and disintegrating rocky bodies around
both M-dwarfs (e.g. K2-22; \citealt{2014ApJ...784...40R}) and white
dwarfs (e.g. WD~1145; \citealt{2015Natur.526..546V}).

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show nearly periodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle.  Most CQVs
are pre-main-sequence M-dwarfs without near-infrared excesses, with
ages of $\approx$5-150 million years (Myr), and rotation periods of at
most two days; they are observed to be $\approx$1-3\% of M-dwarfs
younger than 100 Myr \citep{2016AJ....152..114R,2022AJ....163..144G}.
The dips can be chromatic, with a reddening law plausibly consistent
with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with a few
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The more likely
scenarios invoke sharp geometries with material extrinsic to the
stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  In the
``clump'' scenario, opaque dust clumps orbiting near the Keplerian
corotation radius eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  In the
``prominence'' scenario, long-lived condensations of plasma are
trapped along the star’s magnetic field lines, also near corotation
\citep{2022MNRAS.514.5465W}.  There is also a ``screen'' scenario, in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

In our view, the arguments seem best established for the dust clump
hypothesis \citep{2023MNRAS.518.4734S}.  This is primarily because the
prominence scenario lacks a mechanism for generating broadband
opacity, and the screen scenario is inconsistent with the degree of
observed periodicity, the lack of infrared excess, and the observed
lifetime of the effect, which is an order of magnitude longer than the
$\approx$$10^7$ year timescale typically quoted for the primordial
disk to disperse.  However, unambiguous evidence, for instance a
spectroscopic detection of silicates during a ``dip'', has yet to be
acquired.

CQVs have been challenging to understand because they have been both
hard to discover and hard to characterize.   Discoverability is tied
to rarity: CQVs comprise about 1\% of the youngest 1\% of M-dwarfs
\citep{2018AJ....155..196R}.  Out of the millions of stars monitored
by K2 and TESS, about 50 CQVs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
The known CQVs are correspondingly faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  This renders time-series spectroscopy
at high resolution technically impossible, despite its potential
utility in ruling between the models.

One way to help rule between the mechanisms is to therefore find
bright and nearby CQVs, since these objects will be the most amenable
to detailed photometric and spectroscopic analyses.  To do this, in
this work, we use 120-second cadence data acquired by TESS between
July~2018 and Sep~2022 (Sectors 1-55; Cycles 1-4).  We present our
search methods in Section~\ref{sec:methods}; and the properties of the
resulting CQV catalog in Section~\ref{sec:results}.  Open questions
are discussed in Section~\ref{sec:discussion}, and we conclude in
Section~\ref{sec:conclusion}.

A word on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs should not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses whose optical variability is linked
to obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  At the risk of
introducing yet another standard, we prefer a nomenclature that
reflects how, when observed over timescales of more than tens of
cycles, CQVs are almost, but not exactly periodic.  Their shape
evolution implies that they are irregularly periodic, or for short,
quasiperiodic.  While the three-type classification scheme proposed by
\citet{2017AJ....153..152S} may indeed provide some helpful visual
distinctions amongst CQV sub-classes, it seems likely that they are
all explained by a single underlying phenomenon, and so we opt to
refer to them by a single empirically descriptive name.

@@

To assess the completeness of the resulting 120-second cadence data
that is the basis of this study, we cross-matched TIC8 \citep{2018AJ....156..102S}
against the Gaia DR2 point-source catalog \citep{2018A&A...616A...1G}.
We opted for Gaia DR2 rather than DR3 because the base catalog for
TIC8 was Gaia DR2, which facilitated a one-to-one crossmatch using the
Gaia source identifiers.  This exercise showed that for $T$$<$16
M-dwarfs, the TESS 2-minute data are roughly 50\% complete at
$\approx$50\,pc.  At $<$20\,pc, $\gtrsim$80\% of the $T$$<$16 M-dwarfs
have at least one sector of short-cadence data; at $>$100\,pc,
$\lesssim$10\% of such M-dwarfs have at least one sector of
short-cadence data.  Armed with this understanding, we then used our
cross-match between Gaia DR2 and TIC8 to select our stars of interest,
which we defined as stars with 120-second cadence TESS light curves
that satisfied
\begin{align}
  T &< 16 \quad&(\mathrm{Bright\ for\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only}) \\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})  \\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}),
\end{align}
for $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ the Gaia $G$-band
absolute magnitude, $\varpi_{\rm as}$ the parallax in units of
arcseconds, and a geometric distance $d$ defined by inverting the
parallax and ignoring any zero-point correction.  This selection
function includes dwarf stars later than spectral types of
$\approx$K6V \citep{2013ApJS..208....9P}.  The target sample therefore
includes \nstarssearched\ M-dwarfs and late-K dwarfs,
down to $T$$<$16 and out to $d$$<$150\,pc.

@@

Table~\ref{tab:thetable} lists the \ncpvsfound\ CQVs identified by our
search, along important physical and observational properties.
\ngoods\ objects demonstrated what we viewed as unambiguous
characteristics of the CQV phenomenon in at least one TESS sector;
we refer to these as the ``gold sample''.  We found the remaining \nmaybes\ CQV
candidates to be ambiguous; their variability might
be caused by say, multiple starspot groups.  Additional data from TESS
or other instruments could help resolve their classification.  In the
following discussion, we will primarily restrict our discussion to the 
gold sample.

@@

The two ``good'' IR excesses belonged to ``candidate'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labelling,
we labelled both as ``maybe'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labelling them, inspection
of further sectors clarified that both sources are dippers (see plots
in Appendix~\ref{app:vetting}).  In addition, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labelled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It therefore also a dipper that snuck its way into our
``maybe'' pile.

@@

Comparing these results against the backdrop of our increasing
understanding of dippers
\citep[e.g.][]{2014AJ....147...82C,2016ApJ...816...69A,2021ApJ...908...16R,2022ApJS..263...14C},
it is clear that the loss of an infrared excess is associated with
strong changes in a star's optical variability.  It is reasonable to
imagine connections between CQVs and dippers: both classes of object
can show transient flux dips that are relatively narrow in duration.
Such dips are probably caused by clumps of dust.  However the CQV dips
are typically {\it i)} more periodic and {\it ii)} less deep than
those of dippers, and {\it iii}) they display far less transience over
timescales of a few to tens of cycles.  This is probably because CQV
stars have demonstrably less dust than dipper stars.  At a population
level, the CQV stars are also older.  A common mystery between the
CQVs and dippers, in our own estimation, is how exactly the {\it
narrowness} of the dust clumps is produced.  It is not unreasonable to
imagine a similar mechanism operating for both types of object, tied
perhaps to a shared magnetic topology, or perhaps to a preference for
dust to inspiral to the star in clumped structures.

\section{CQV evolution}

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CQVs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CQVs in our catalog; for plotting purposes we show
the brightest 27.
Of these 27 CQVs, a few show clear signs of the phenomenon in one
sector, and marginal or non-existence signs in the other.
While there is some subjectivity in this assessment, to our eyes cases
for which at least one sector would be flagged as ``ambiguous''
include
TIC~368129164 (Sector 23 might be labelled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labelled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
implicitly have already given this source such a label!  Based on the
the shape evolution between Sectors 13 and 39 however, it is a CQV.
One could easily also imagine in cases like TIC~206544316 that if
observed at lower signal to noise, the drastic shape evolution would
not be appreciated.  We emphasize that the periods themselves were all
stable to $<$0.1\%.  Broadly speaking, this shape evolution suggests
that CQVs have an ``on/off'' duty cycle of $\approx$75\%.  This type
of correction is likely worth including in population-level estimates
of how intrinsically common CQVs are in the low-mass stellar
population \citep[e.g.][]{2022AJ....163..144G}.

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{TIC_402980664_P18.5611_2min_phase_timegroups_mosaic_ymin-4.8_ymax3.pdf}
		\vspace{-0.45cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      time gaps; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is very sharp: it shows a flux
exercusion of 3.5\% over about 22 minutes (0.02\,$P$), which is the
steepest slope exhibited anywhere in this object's remarkable dataset.
After the sharp dip, there is a roughly exponential fall-off spanning
about a quarter of a period, punctuated by coherent local minima and
maxima which in detail (Appendix~\ref{fig:app:lp}) have slightly
longer periods than the sharp dip.  The sharp leading dip only
decreases during a sudden state-switch at BTJD 2030.7 (cycles
299-315), which happens immediately during a flare
(Figure~\ref{fig:lplc}).  The trailing dips remain afterwards.

@@

{\sc Dips grow more slowly than they disappear}---Although there are a
few instances in which we observe dips ``switch off'' over less
than one cycle, dip growth seems to happen more slowly.  For instance,
the dip that grows between phase 0 and 0.1 between cycles 260-290
begins to become visible around BTJD 1993.2, and growths in depth by
about 2\% over about six cycles, to become easily detectable by eye by
BTJD 1997.7.  The evolution of this particular dip is most clear in
the river plots.  The evolution of the latter dip group in cycles
1410-1481 is another example of this slow mode of dip growth.

@@

\subsection{Extremes}
The closest CQV in our catalog is DG~CVn (TIC~368129164;
$d$$=$18\,pc), which is a likely member of AB Dor.  To our knowledge,
this manuscript is the first time that it has been noted as a CQV.
The three brightest CQVs we have found are DG~CVn ($T$=9.3),
TIC~405754448 ($T$=9.6), and TIC~167664935 ($T$=10.3).

The shortest period belongs to TIC~201789285, at 3.64\,hr.
The longest period belongs to TIC~405910546, at 37.9\,hr.  If
that source turns out to be an eclipsing binary, then the next-longest
would be TIC~193831684 (31.0\,hr).

The lowest mass ($\approx 0.12$\,M$_\odot$) belongs to TIC~267953787.
The catalog contains a few other stars with similar mass, which
suggests the phenomenon may extend into the brown dwarf regime.

>>>>

@@

%\title{Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS}
%\title{Complex Quasiperiodic Variables From Four Years of TESS}
\title{Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS}

@@

% primary authors
\author{Author list and order not final!}
\author[0000-0002-0514-5538]{Luke~G.~Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{\caltech}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust or gas clumps in orbit at the
  Keplerian corotation radius.  Here, we report new CQVs discovered in
  TESS data collected between July~2018 and Sep~2022.
  Our search of \nstarssearched\ K- and M-dwarfs
  with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\ gold-standard CQVs.
  Most of these discoveries are new, and they include the brightest
  and closest examples of this class of object known ($T$$\approx$9.5;
  $d$$\approx$20\,pc), as well as the most massive
  ($\approx$0.65\,M$_\odot$).  A few objects are outliers among
  outliers; LP 12-502 for instance shows a ``dip complex'' with a
  period and total duration that are fixed over more than 1{,}000 cycles,
  but which in detail exhibits anywhere from four to eight local minima
  per cycle.  LP-502 also displayed distinct superposed periods
	simultaneously, and displayed drastic shape changes shortly after flares.
	Broadly speaking, we find that none of the CQVs maintain a fixed light curve
	shape over timescales of more than a few hundred cycles, and we revisit the
	arguments for why transient corotating material is the most viable explanation.
	In the future, we expect that our sample will facilitate modeling and
	observational efforts aimed at understanding these objects, and connecting them
	to the broader contexts of star, disk, and exoplanet evolution.
	For instance, if the transiting material is dust, then the rate of dip evolution
	implies that of order an asteroid belt's worth of mass passes through these
	structures over the $10^8$ years for which they are observed.
\end{abstract}

@@

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.  More exotic forms of optical variability
relevant to this work include transiting exocomets (e.g. $\beta$~Pic;
\citealt{2019A&A...625L..13Z}) and disintegrating rocky bodies around
both M-dwarfs (e.g. KOI-2700; \citealt{2014ApJ...784...40R}) and white
dwarfs (e.g. WD~1145; \citealt{2015Natur.526..546V}).

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show nearly periodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle.  Most 
are pre-main-sequence M-dwarfs without near-infrared excesses, 
ages of $\approx$5-150 million years (Myr), and rotation periods of at
most two days; they are observed to be $\approx$1-3\% of M-dwarfs
younger than 100 Myr \citep{2016AJ....152..114R,2022AJ....163..144G}.
The dips can be chromatic, with a reddening law plausibly consistent
with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with some
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The timescales and
amplitudes of the flux variations instead require sharp geometries
with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  The
``clump'' scenario invokes opaque dust clumps orbiting near the Keplerian
corotation radius, which periodically eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The
``prominence'' scenario invokes long-lived condensations of cool and dense
marginally-ionized gas, embedded within the
hotter corona, that would be centrifugally supported near corotation
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
Such structures are analogous to quiescent prominences and filaments
observed on the Sun, though at much larger relative distances from the
stellar surface.
A final posibility that has been suggested is one of a ``screen'', in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

In our view, the dust clump and prominence hypotheses seem most
plausible.  They are qualitatively similar, except one
invokes opacity from dust, while the other invokes opacity from gas,
likely bound-free transitions in hydrogen or perhaps a
molecular opacity.  The screen scenario seems inconsistent with the
degree of observed periodicity, the lack of infrared excess, and the
observed lifetime of CQVs extending an order of magnitude longer than
the $\approx$$10^7$ year timescale typically quoted for the primordial
disk to disperse.  However, unambiguous evidence 
has yet to be acquired in support of any scenario.  Such evidence
might include a spectroscopic detection of silicate 10\,$\mu$m dust
absorption during a dip, or perhaps detection of transient
Balmer-line absorption during the dips, similar to observations made in
systems such as AB~Dor \citep[see the review
by][]{1999ASPC..158..146C}.

CQVs remain mysterious because they have been both hard to discover
and hard to characterize.   Discoverability is tied to rarity: CQVs
comprise about 1\% of the youngest 1\% of M-dwarfs
\citep{2018AJ....155..196R}.  Out of the millions of stars monitored
by K2 and TESS, about 50 CQVs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
The known CQVs are correspondingly faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  This renders time-series spectroscopy
at high resolution technically impossible, despite its potential
utility in ruling between the models.

One way to help rule between the mechanisms is to therefore find
bright and nearby CQVs, since these objects will be the most amenable
to detailed photometric and spectroscopic analyses.  To do this, in
this work, we use 120-second cadence data acquired by TESS between
July~2018 and Sep~2022 (Sectors 1-55; Cycles 1-4).  We present our
search methods in Section~\ref{sec:methods}, and the properties of the
resulting CQV catalog in Section~\ref{sec:results}.  
The evolution of many CQVs over a two-year baseline is described
in Section~\ref{sec:evoln}, including a deep-dive into LP 12-502.
Implications 
are discussed in Section~\ref{sec:discussion}, and we conclude in
Section~\ref{sec:conclusion}.

A word on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs should not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses whose optical variability is linked
to obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  At the risk of
introducing yet another standard, we prefer a nomenclature that
reflects how, when observed over timescales of more than tens of
cycles, CQVs are almost but not exactly periodic.  They are
irregularly periodic, or for short, quasiperiodic.  While the
three-type classification scheme proposed by
\citet{2017AJ....153..152S} may indeed provide some helpful visual
distinctions amongst CQV sub-classes, it seems likely that they are
all explained by a single underlying phenomenon, and so we opt to
refer to them by a single empirically descriptive name.

@@

To assess the completeness of the resulting 120-second cadence data
that is the basis of this study, we cross-matched TIC8
\citep{2018AJ....156..102S} against the Gaia DR2 point-source catalog
\citep{2018A&A...616A...1G}.  We opted for Gaia DR2 rather than DR3
because the base catalog for TIC8 was Gaia DR2, which facilitated a
one-to-one crossmatch using the Gaia source identifiers.  This
exercise showed that for $T$$<$16 M-dwarfs, the TESS 2-minute data are
roughly 50\% complete at $\approx$50\,pc.  At $<$20\,pc, $\gtrsim$80\%
of the $T$$<$16 M-dwarfs have at least one sector of short-cadence
data; at $>$100\,pc, $\lesssim$10\% of such M-dwarfs have at least one
sector of short-cadence data.  Armed with this understanding, we then
used our cross-match between Gaia DR2 and TIC8 to select our stars of
interest, which we defined as stars with 120-second cadence TESS light
curves that satisfied
\begin{align}
  T &< 16 \quad&(\mathrm{Bright\ for\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only}) \\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})  \\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}),
\end{align}
for $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ the Gaia $G$-band
absolute magnitude, $\varpi_{\rm as}$ the parallax in units of
arcseconds, and a geometric distance $d$ defined by inverting the
parallax and ignoring any zero-point correction.  This selection
function includes dwarf stars later than spectral types of
$\approx$K6V \citep{2013ApJS..208....9P} for which TESS can acquire
1\% relative precision photometry in 1 hour of observation
\citep{2015JATIS...1a4003R}.  The target sample therefore includes
\nstarssearched\ M-dwarfs and late-K dwarfs, down to $T$$<$16 and out
to $d$$<$150\,pc.

@@

Table~\ref{tab:thetable} lists the \ncpvsfound\ CQVs identified by our
search, along with important physical and observational properties.
\ngoods\ objects demonstrated what we viewed as unambiguous
characteristics of the CQV phenomenon in at least one TESS sector; we
refer to these as the ``gold sample''.  We found the remaining
\nmaybes\ CQV candidates to be ambiguous; their variability might be
caused by say, multiple starspot groups.  Additional data from TESS or
other instruments could help resolve their classification.  In the
following discussion, we will primarily restrict our discussion to the
gold sample.

@@

The two ``good'' IR excesses belonged to ``candidate'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labelling,
we labelled both as ``maybe'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labelling them, inspection
of further sectors clarified that both sources are dippers (see plots
in Appendix~\ref{app:vetting}).  In addition, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labelled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It therefore also a dipper that snuck its way into our
``maybe'' pile.
Possible evolutionary connections between CQVs and dippers are
highlighted in Section~\ref{subsec:discdippers}.

@@

\section{CQV evolution}
\label{sec:evoln}

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CQVs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CQVs in our catalog; for plotting purposes we show
the brightest 27.
Of these 27 CQVs, a few show clear signs of the phenomenon in one
sector, and marginal or non-existence signs in the other.
While there is some subjectivity in this assessment, to our eyes cases
for which at least one sector would be flagged as ``ambiguous''
include
TIC~368129164 (Sector 23 might be labelled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labelled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
implicitly have already given this source such a label.  However, based on the
the shape evolution between Sectors 13 and 39, it is a CQV.
One could easily also imagine in cases like TIC~206544316 that if
observed at lower signal to noise, the drastic shape evolution would
not be appreciated.  We emphasize that the periods themselves were all
stable to $<$0.1\%.  Broadly speaking, this shape evolution suggests
that CQVs have an ``on/off'' duty cycle of $\approx$75\%.  This type
of correction is likely worth including in population-level estimates
of how intrinsically common CQVs are in the low-mass stellar
population \citep[e.g.][]{2022AJ....163..144G}.

@@

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f5.pdf}
		\vspace{-0.45cm}
		\caption{
      {\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
      epoch over three years.  Each panel shows one (stacked) TESS
      orbit; small text denotes relative cycle number.  There are 200
      binned black points per cycle.  The TESS pointing law dictates
      time gaps; larger gaps tend to yield larger shape changes.  The
      dips usually evolve over tens to hundreds of cycles.  However
      cycles 1233-1264 show a dip that ``switched'' from a depth and
      duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
      cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is very sharp: it shows a flux
exercusion of 3.5\% over about 22 minutes (0.02\,$P$), which is the
steepest slope exhibited anywhere in this object's remarkable dataset.
After the sharp dip, there is a roughly exponential fall-off spanning
about a quarter of a period, punctuated by coherent local minima and
maxima which in detail (Appendix~\ref{app:lp}) have slightly
longer periods than the sharp dip.  The sharp leading dip only
decreases during a sudden state-switch at BTJD 2030.7 (cycles
299-315), which happens immediately during a flare
(Figure~\ref{fig:lplc}).  The trailing dips remain afterwards.

@@

{\sc Slow growth; rapid death}---Although there are a
few instances in which we observe dips ``switch off'' over less
than one cycle, dip growth seems to happen more slowly.  For instance,
the dip that grows between phase 0 and 0.1 between cycles 260-290
begins to become visible around BTJD 1993.2, and growths in depth by
about 2\% over about six cycles, to become easily detectable by eye by
BTJD 1997.7.  The evolution of this particular dip is most clear in
the river plots.  The evolution of the latter dip group in cycles
1410-1481 is another example of this slow mode of dip growth.

@@

\subsection{Extremes}
The closest CQV in our catalog is DG~CVn (TIC~368129164;
$d$$=$18\,pc), a member of AB Dor.  To our knowledge,
this manuscript is the first time that it has been noted as a CQV.
The three brightest CQVs are DG~CVn ($T$=9.3),
TIC~405754448 ($T$=9.6), and TIC~167664935 ($T$=10.3).
The shortest period belongs to TIC~201789285, at 3.64\,hr.
The longest period belongs to TIC~405910546, at 37.9\,hr.  If
the latter source turns out to be an eclipsing binary, the next-longest
would be TIC~193831684 (31.0\,hr).

The lowest mass ($\approx 0.12$\,M$_\odot$) belongs to TIC~267953787.
The catalog contains a few other stars with similar mass.
Given the small number of sub-stellar mass objects in our target sample,
this suggests that future studies of brown dwarf
photometric variability might also yield complex quasiperiodic
variables, though there would be degeneracies in interpretation with planetary surface features such
as clouds and latitudinal bands \citep[e.g.][]{2021ApJ...906...64A,2022ApJ...924...68V}.
